A Hitchhiker's Guide to Reproducible Research" # replace with something serious

### 2021-05-09 manuscript.Rmd
put each sentance on its own line #50

<!--Keep notes.-->
 A great tool to capture ideas or meeting notes is Markdown. Markdown is a lightweight text format that is very close to plain text with a minimal subset of reserved symbols for formatting instructions. This way, Markup does not need any specialized software for editting, it is highly fool-proof (unlike for example LaTeX) and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation. R Markdown is the smart combination of Markdown with dedicated code chunks that are evaluated to formatted text, statistical results, or images when the document is rendered to its final format. We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts and presentations -- and even your CV.
To create a Markdown document, click on:

@@

Preregistration is the specification of a research plan in advance of data analysis. The goal of preregistration is transparency and a clear-cut divide into exploration and confirmation.
Thus, it serves as a tool that may help to increase the credibility of empirical results. <!-- AP citation needed-->
Without getting into too much detail of the meta-scientific arguments for preregistration, we want to note two advantages. <!-- AP can we add a citation to someone who has the arguments?-->
First, a preregistration is an opportunity to plan a study carefully.
Gathering empirical data requires many resources, and better planning increases the chances that these efforts are worthwhile.
For us, a good preregistration is the same as a good plan; the only difference is that a preregistration is made public beforehand.

@@

Currently, there are various templates available that guide researchers in developing data analysis plans for preregistration <!--AP citation needed-->. Still, with any natural-language representation of an analysis plan, there remains ambiguity when turning the natural language description of a given data analysis into code <!--AP are the classic computer science literature who talks about this ambiguity... theoretical CS? NLP research or such?-->. Here, we discuss preregistration as code (PAC) with the R package repro, which leaves no room for ambiguity. PAC essentially reverses the process of empirical research. Before researchers gather data, they write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results --- initially based on simulated data. Then, the source code of this manuscript serves as an unambiguous description of the analysis plan and is published as a preregistration.

@@

One way to ensure that you can expect your analysis to work with real data is to simulate data.
A data simulation allows you to verify that the code runs error-free and lets you form more realistic expectations about the estimates' correctness and precision. <!--AP shouldn't this be in the simulation part?-->
We, for example, simulated data for the Machiavellism subscale of the short dark triad (SD3) and analyzed this with structural equation modelling (SEM).
The simulation showed that we have to gather at least `missing_rcode_simulation` observations per group to archive enough precision and that our solution to a data problem we expect (insufficient measurement invariance) introduces a bias of `missing_rcode_simulation`%.
If you went to the trouble to simulate data, we argue that you get a preregistration almost for free.
A preregistration in code describes sufficiently how you want to analyze the data, but ideally, you also convey the why.
Dynamic document generation combines code and text, enabling you to explain your reasoning.
See the section on [R Markdown] for an example of a dynamic document generation paradigm.

@@

One of the most significant benefits of open research is that other people can freely reuse materials, ideally including everything from stimuli, over computerized experiments, to data analysis code.
To use your materials, they first have to find them and learn how to use them.
Interested users find out about your project via the readme.
Historically[^readme-age], a readme file is the default place to look for human readable information in a computer archive 
the first thing they see when they visit your project on [GitHub].
It is supposed to explain to visitors what your project is about and how to reproduce it.
You could, for example, include the abstract of your (planned) paper here.<!-- AP: shouldn't this be a function in repro? to auto-insert the abstract?-->

@@

<!-- AP: Can we show how to set MIT license for code an CCBY license for the manuscript? Can we make a repro function that creates such a double-license? How to do it actually? Two license files? -->

@@

<!--AP cite WORCS and some arguments for documenting your data. Also, cite the codebook package -->

@@

<!--Hands on: Introduce GitHub PR and review-->
There are various ways, in which other scholars may interact with your published repository. For example, in our exposition of the workflow for reproducibility [@Peikert2019], once we had published our preprint, scholars previously unknown to us started to send us spelling corrections. When multiple persons start modifying a project in parallel, keeping track of the edits and how they are supposed to be merged is crucial. Clearly, allowing everyone to edit the project is difficult <!--AP are there papers who analyzed Wikipedia and the way concurrent changes and discussions are resolved - is this relevant at all?-->

@@

<!-- AP: this is a nice example but why is this an abstract example now? I thought that 
we walk through one(!) example. I think this really should be the concrete Makefile from
the analysis that we are talking about -->

@@

Modern data analysis is only feasible because researchers rely on much external software. <!-- AP: can we give examples of cool applications in psychology -->
One downside is that a considerable number of external dependencies may hinder collaboration and long term reproducibility. The dependency on a single R package may easily create a dependency tree of more than hundred direct or indirect dependencies [@epskamp2019].
The critical problem is the fragility of a software setup.
Many components have to fit perfectly together, and each change, e.g. an update, risks that balance . At the same time, software updates are very much necessary to remove errors or improve methodology [@epskamp2019].
Another problem is that the software environment is also variable and what works on one machine might not work on any other.
Containers directly address such issues.
Instead of leaving it to the user which software should be installed how on their computer, Docker provides all software used in an analysis, including the operating system itself. The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such. <!--AP can we add an example here with the non-reproducibility of an eigen problem solver - I think we discussed at some point in time when WOlfgang Viechtbauer sent us a link to a paper -->.
Docker does this without interfering with the already installed software by using a virtual software environment that is independent of the host software environment.
Such a snapshot of the software stack is called an "image".
Packaging all needed software in such an image requires considerable amounts of storage space.
There are two major strategies helping to keep the storage requirements reasonable.
First, there is a community that maintains images for particular purposes.
For example, there are ready-made images for R based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they additionally need on top of these pre-made images.
The image that was used for this article uses `r round(as.numeric(fs::file_size("reprotutorial.sif"))/1024^3, 2)`Gb of disc space.
It does include Ubuntu, R, RStudio, LaTeX as well as many R packages like the tidyverse [@tidyverse], amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.
Second, instead of saving a binary image of the software, one can also save a container recipe, a so-called  `Dockerfile`. This recipe is a textual description of all commands that need to be executed to recreate the image.
Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and are an unambiguous description of how to recreate a given container image. However, they rely on the assumption that all software repositories will remain accessible and will continue to make available historic software versions.
For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).
With the `repro`-package adding R packages dependencies is as simple as adding them to the header under the tag `packages:`

```
---
title: "Preregistration"
author: "Aaron Peikert"
date: "4/12/2021"
output: html_document
repro:
  scripts:
    - R/simulation.R
  packages:
    - tidyverse
    - here
    - lavaan
    - furrr
---
```

With the R version that you currently use, this specification is then translated into a `Dockerfile` that makes sure that the container image has all these packages installed.
The `Dockerfile` also contains the current date and fixes packages versions to the version of the package that were available on that day on the [Comprehensive R Archive Network (CRAN)](https://cran.r-project.org)[^cran] by relying on the [CRAN timemachine of Microsoft R Application Network (MRAN)](https://mran.microsoft.com/timemachine)[^mran].
<!-- Hands on: Introduce Docker -->

@@

Docker can be a bit unfriendly, because it requires a complicated mental model and some knowledge about how to install software on Linux.
An R-focused alternative is the R package [`renv`](https://rstudio.github.io/renv/articles/renv.html)[^renv][@R-renv].
`renv` isolates all R packages used in a project and records there version as well as the source from where they where installed.
This mechanism is faster and simpler to use than Docker, at the expense of only tracking R packages.
`renv` therefore is already of great use for reproducibility.
However, even when your project relies solely on R, external dependencies can still change the outcome of computations, like the operating system, the compilers used, the BLAS system in use and so forth.
Most importantly the R version itself has to match.

@@

As default, `repro` uses Docker, which is the defacto standard for cloud computing, as a container solution.
However, Docker requires administrator rights for execution. Whereas this is no issue on most modern cloud computing infrastructures, it is often an issue in classic HPC environments.
Singularity <!--AP add reference --> is an alternative solution. It was designed to run without administratot rights, while maintaining full compatibility with Docker images.
If you already have a Docker image, you can choose if you want to run this image with Docker or with Singularity.
Importantly, this choice does not affect how code is executed inside the container.
Like with Docker, the default `Makefile` generated by `repro` comes with a mechanism to automatically send all commands to Singularity. To reproduce an analysis with Singularity, simple add the parameter `SINGULARITY=TRUE` as follows:

@@

Note that both this tutorial as well as our article on the reproducible workflow [@Peikert2019] were created using this workflow and can be reproduced from their source code repositories <!--AP add links -->.
<!-- maybe we add here the real data? #25 -->

@@

<!--AP: make sure to cite WORCS and describe what WORCS is proposing here-->
<!-- Hands-on: check changed hash -->
A prerequisite to reproducibility is the availability of code and data <!-- AP: can you add the reference to this tight reproducibility requirements that you told me about?-->. We use containers to guarantee that all code components are available and work as expected. Ideally, all data is also stored in the repository. In practice, however, there are various reasons, which may prohibit you from making your data available, for example legal or regulatory or privacy concerns.
You may find yourself in a situation where you can make the data only available on request. <!-- AP please cite the SHARE data and explain as an example what one has to do to ontain these data -->.
A major threat to reproducibility is that the data has changed in the meantime or multiple versions are circulating and it is not exactly clear what version was used originally.
To at least check that the data was not changed, we suggest to provide a so called "hash", which is a cryptographic fingerprint.
If the data changes, so does its fingerprint or hash.
The user can therefore compare if the data they have received is the same that was used in the original publication, as long as the original publication does contain the hash of the data.

To verify the data integrity, you can include a line of code after reading the data that looks like this:
<!-- AP: shouldn't this be wrapped into a repro function and also stored in the YAML header??? -->

@@

Here, we provided a step-by-step tutorial with what we think are best practices for creating a fully reproducible repository for data analysis projects that center on the statistical programming language R. 
The workflow avoids common threats to reproducibility, such as copy-and-paste errors, missing files, missing software dependencies, or side-effects through changing dependent software packages, changes in the programming language, or even at the level of the operating system itself. 
However, the workflow as proposed by @Peikert2019 requires mastering up to four different software tools and their specification languages, which are typically not taught in the curriculum of social and behavioral sciences but rather belong to computer science and software development. 
Currently, this is a serious obstacle in adopting the workflow for researchers who strive for reproducibility but have little training in computer science and software development.  
The repro package, which we introduced here, supports scholars in creating reproducible workflows. It abstracts tedious and error-prone tasks such as the creation of a Makefile or the specification of a container.
By extending the YAML specification that is used for storing meta information in R Markdown files, which are becoming increasingly popular among R users, repro helps users to implement @Peikert2019's workflow with minimal effort. Most importantly, repro is implemented such that the package itself is not needed to reproduce a workflow that was generated according to the workflow. 
Importantly, this means that users that adopt the workflow do not have to rely on the future availability and maintenance of the repro package. 
<!-- AP: add more. Summarize. Talk about the benefits. Talk about the limitations. What are the threats to reproducibility. Make the case that adopting this workflow is easy and can even be done for existing repositories. Also, wasn't an important point of repro that it itself is not a dependency? I think this needs to be strengthened and probably already said in the intro-->

>>>>

<!--Keep notes.-->
 A great tool to capture ideas or meeting notes is Markdown.
 Markdown is a lightweight text format that is very close to plain text with a minimal subset of reserved symbols for formatting instructions.
 This way, Markup does not need any specialized software for editing, it is highly fool-proof (unlike for example LaTeX) and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
 R Markdown is the smart combination of Markdown with dedicated code chunks that are evaluated to formatted text, statistical results, or images when the document is rendered to its final format
 We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts and presentations---and even your CV.
To create a Markdown document, click on:

@@

Preregistration is the specification of a research plan in advance of data analysis.
The goal of preregistration is transparency and a clear-cut divide into exploration and confirmation.
Thus, it serves as a tool that may help to increase the credibility of empirical results.
Without getting into too much detail of the meta-scientific arguments for preregistration, we want to note two advantages.
First, a preregistration is an opportunity to plan a study carefully.
Gathering empirical data requires many resources, and better planning increases the chances that these efforts are worthwhile.
For us, a good preregistration is the same as a good plan; the only difference is that a preregistration is made public beforehand.

@@

Currently, there are various templates available that guide researchers in developing data analysis plans for preregistration. 
Still, with any natural-language representation of an analysis plan, there remains ambiguity when turning the natural language description of a given data analysis into code.
Here, we discuss preregistration as code (PAC) with the R package repro, which leaves no room for ambiguity.
PAC essentially reverses the process of empirical research.
Before researchers gather data, they write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results --- initially based on simulated data.
Then, the source code of this manuscript serves as an unambiguous description of the analysis plan and is published as a preregistration.

@@

One way to ensure that you can expect your analysis to work with real data is to simulate data.
A data simulation allows you to verify that the code runs error-free and lets you form more realistic expectations about the estimates' correctness and precision.
We, for example, simulated data for the Machiavellism subscale of the short dark triad (SD3) and analyzed this with structural equation modelling (SEM).
The simulation showed that we have to gather at least `missing_rcode_simulation` observations per group to archive enough precision and that our solution to a data problem we expect (insufficient measurement invariance) introduces a bias of `missing_rcode_simulation`%.
If you went to the trouble to simulate data, we argue that you get a preregistration almost for free.
A preregistration in code describes sufficiently how you want to analyze the data, but ideally, you also convey the why.
Dynamic document generation combines code and text, enabling you to explain your reasoning.
See the section on [R Markdown] for an example of a dynamic document generation paradigm.

@@

One of the most significant benefits of open research is that other people can freely reuse materials, ideally including everything from stimuli, over computerized experiments, to data analysis code.
To use your materials, they first have to find them and learn how to use them.
Interested users find out about your project via the readme.
Historically[^readme-age], a readme file is the default place to look for human readable information in a computer archive 
the first thing they see when they visit your project on [GitHub].
It is supposed to explain to visitors what your project is about and how to reproduce it.
You could, for example, include the abstract of your (planned) paper here.

@@

@@

@@

<!--Hands on: Introduce GitHub PR and review-->
There are various ways, in which other scholars may interact with your published repository.
For example, in our exposition of the workflow for reproducibility [@Peikert2019], once we had published our preprint, scholars previously unknown to us started to send us spelling corrections.
When multiple persons start modifying a project in parallel, keeping track of the edits and how they are supposed to be merged is crucial.
Clearly, allowing everyone to edit the project is difficult.

@@

@@

Modern data analysis is only feasible because researchers rely on much external software.
One downside is that a considerable number of external dependencies may hinder collaboration and long term reproducibility.
The dependency on a single R package may easily create a dependency tree of more than hundred direct or indirect dependencies [@epskamp2019].
The critical problem is the fragility of a software setup.
Many components have to fit perfectly together, and each change, e.g. an update, risks that balance .
At the same time, software updates are very much necessary to remove errors or improve methodology [@epskamp2019].
Another problem is that the software environment is also variable and what works on one machine might not work on any other.
Containers directly address such issues.
Instead of leaving it to the user which software should be installed how on their computer, Docker provides all software used in an analysis, including the operating system itself.
The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
Docker does this without interfering with the already installed software by using a virtual software environment that is independent of the host software environment.
Such a snapshot of the software stack is called an "image".
Packaging all needed software in such an image requires considerable amounts of storage space.
There are two major strategies helping to keep the storage requirements reasonable.
First, there is a community that maintains images for particular purposes.
For example, there are ready-made images for R based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they additionally need on top of these pre-made images.
The image that was used for this article uses `r round(as.numeric(fs::file_size("reprotutorial.sif"))/1024^3, 2)`Gb of disc space.
It does include Ubuntu, R, RStudio, LaTeX as well as many R packages like the tidyverse [@tidyverse], amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.
Second, instead of saving a binary image of the software, one can also save a container recipe, a so-called  `Dockerfile`.
This recipe is a textual description of all commands that need to be executed to recreate the image.
Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and are an unambiguous description of how to recreate a given container image.
However, they rely on the assumption that all software repositories will remain accessible and will continue to make available historic software versions.
For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).
With the `repro`-package adding R packages dependencies is as simple as adding them to the header under the tag `packages:`

```
---
title: "Preregistration"
author: "Aaron Peikert"
date: "4/12/2021"
output: html_document
repro:
  scripts:
    - R/simulation.R
  packages:
    - tidyverse
    - here
    - lavaan
    - furrr
---
```
<!--AP emphasize more how cool it is that this stuff is automatically generated by repro!-->
With the R version that you currently use, this specification is then translated into a `Dockerfile` that makes sure that the container image has all these packages installed.
The `Dockerfile` also contains the current date and fixes packages versions to the version of the package that were available on that day on the [Comprehensive R Archive Network (CRAN)](https://cran.r-project.org)[^cran] by relying on the [CRAN timemachine of Microsoft R Application Network (MRAN)](https://mran.microsoft.com/timemachine)[^mran].
<!-- Hands on: Introduce Docker -->

@@

Docker can be a bit unfriendly, because it requires a complicated mental model and some knowledge about how to install software on Linux.
An R-focused alternative is the R package [`renv`](https://rstudio.github.io/renv/articles/renv.html)[^renv][@R-renv].
`renv` isolates all R packages used in a project and records there version as well as the source from where they where installed.
This mechanism is faster and simpler to use than Docker, at the expense of only tracking R packages.
`renv` therefore is already of great use for reproducibility [@vanlissa2020worcs].
However, even when your project relies solely on R, external dependencies can still change the outcome of computations, like the operating system, the compilers used, the BLAS system in use and so forth.
Most importantly the R version itself has to match.
For example, @Peikert2019 illustrate how the generation of random numbers across different R versions is a threat to reproducibility.

@@

As default, `repro` uses Docker, which is the defacto standard for cloud computing, as a container solution.
However, Docker requires administrator rights for execution.
Whereas this is no issue on most modern cloud computing infrastructures, it is often an issue in classic HPC environments.
Singularity is an alternative solution.
It was designed to run without administratot rights, while maintaining full compatibility with Docker images.
If you already have a Docker image, you can choose if you want to run this image with Docker or with Singularity.
Importantly, this choice does not affect how code is executed inside the container.
Like with Docker, the default `Makefile` generated by `repro` comes with a mechanism to automatically send all commands to Singularity.
To reproduce an analysis with Singularity, simple add the parameter `SINGULARITY=TRUE` as follows:

@@

Note that both this tutorial as well as our article on the reproducible workflow [@Peikert2019] were created using this workflow and can be reproduced from their source code repositories.
<!-- maybe we add here the real data? #25 -->

@@

<!-- Hands-on: check changed hash -->
A prerequisite to reproducibility is the availability of code and data.
We use containers to guarantee that all code components are available and work as expected.
Ideally, all data is also stored in the repository.
In practice, however, there are various reasons, which may prohibit you from making your data available, for example legal or regulatory or privacy concerns.
You may find yourself in a situation where you can make the data only available on request.
A major threat to reproducibility is that the data has changed in the meantime or multiple versions are circulating and it is not exactly clear what version was used originally.
To at least check that the data was not changed, we suggest to provide a so called "hash", which is a cryptographic fingerprint.
If the data changes, so does its fingerprint or hash.
The user can therefore compare if the data they have received is the same that was used in the original publication, as long as the original publication does contain the hash of the data.

To verify the data integrity, you can include a line of code after reading the data that looks like this:

@@

Here, we provided a step-by-step tutorial with what we think are best practices for creating a fully reproducible repository for data analysis projects that center on the statistical programming language R.
The workflow avoids common threats to reproducibility, such as copy-and-paste errors, missing files, missing software dependencies, or side-effects through changing dependent software packages, changes in the programming language, or even at the level of the operating system itself.
However, the workflow as proposed by @Peikert2019 requires mastering up to four different software tools and their specification languages, which are typically not taught in the curriculum of social and behavioral sciences but rather belong to computer science and software development.
Currently, this is a serious obstacle in adopting the workflow for researchers who strive for reproducibility but have little training in computer science and software development.
The repro package, which we introduced here, supports scholars in creating reproducible workflows.
It abstracts tedious and error-prone tasks such as the creation of a Makefile or the specification of a container.
By extending the YAML specification that is used for storing meta information in R Markdown files, which are becoming increasingly popular among R users, repro helps users to implement @Peikert2019's workflow with minimal effort.
Most importantly, repro is implemented such that the package itself is not needed to reproduce a workflow that was generated according to the workflow.
Importantly, this means that users that adopt the workflow do not have to rely on the future availability and maintenance of the repro package.
